\subsection{Increased flexibility for trace file contents}

The trace files produced by the current version of the VICE VDMTools
are not in a format that is appropriate for easy post-analysis because
a proprietary ASCII format is used. The ``ShowVICE'' tool thus has to
parse these trace events before they can be analysed. This
can be solved if the ShowVICE functionality is built into VDMTools as
suggested above. But, the contents of each trace event are still quite
limited, e.g. the arguments for operations invoked are not presented in
the trace file. The reason for that is simple and pragmatic; it would
slow down the execution of the interpreter too much if all information
available inside the interpreter was dumped with each trace event.

Thus, the ideal solution would be to enable the user to select which
information (s)he would like to trace during the simulation and under which
circumstances it should do so. This could, for example, be implemented with
a small scripting language enabling the user to configure the type and the
amount of data to be logged in the trace file and thus be available for
subsequent post-analysis (as a VDM value instead of a proprietary ASCII
format). Some of the timing requirements for an embedded real-time system
may otherwise not be expressible without this feature. An earlier draft of
this note has already led to a significant improvement of VICE. The
\verb+time+ keyword was added to the language, such that simulation
wall clock is accessible from within the specification.


